Discrete Probability for Discrete Quantum Computing

### 2016-05-03 prob.tex
small changes to first page

@@

\begin{example}[Two-coin probability space]\label{ex1}
  Consider an experiment that tosses two coins. We have four possible
  outcomes that constitute the sample space
  $\Omega = \{ HH, HT, TH, TT \}$. There are 16 total events including
  for example the event $\{ HH, HT \}$ that the first coin lands heads
  up, the event $\{ HT, TH \}$ that the two coins land on opposite
  sides, and the event $\{ HT, TH, TT\}$ that at least one coin lands
  tails up. Here is a possible probability measure for these events:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& 0 \\
\pmeas(\{ HH \}) &=& 1/3 \\
\pmeas(\{ HT \}) &=& 0 \\
\pmeas(\{ TH \}) &=& 2/3 \\
\pmeas(\{ TT \}) &=& 0 \\
\pmeas(\{  HH, HT \}) &=& 1/3 \\
\pmeas(\{  HH, TH \}) &=& 1 \\
\pmeas(\{  HH , TT \}) &=& 1/3 
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& 2/3 \\
\pmeas(\{  HT , TT \}) &=& 0 \\
\pmeas(\{  TH , TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH \}) &=& 1 \\
\pmeas(\{  HH, HT, TT \}) &=& 1/3 \\
\pmeas(\{  HH, TH, TT \}) &=& 1 \\
\pmeas(\{  HT, TH, TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH, TT \}) &=& 1
\end{array}
\end{array}\]
The assignment satisfies the two constraints for probability measures:
the probability of the entire sample space is 1, and the probability
of every collection of disjoint events (e.g.,
$\{ HT, TH \} = \{ HT \} \cup \{ TH \}$) is the sum of the individual
probabilities. The probability of collections of non-disjoint events
(e.g., $\{ HT, TH, TT \} = \{ HT, TH \} \cup \{ TH , TT \}$) may add
to something different than the probabilities of the individual
events. It is useful to think that this probability measure is
completely induced by the two coins in question and their
characteristics in the sense that each pair of coins induces a
measure, and each measure must correspond to some pair of coins.  The
measure above is induced by two coins such that the first coin is
twice as likely to land tails up than heads up and the second coin is
double-headed.  
\qed\end{example}

@@

\paragraph*{Set-valued Probability Measures.} The simplest such set is
the set $\mathscr{L}_{2}=\left\{ \imposs, \poss\right\}$ together with
the operation $\vee$ where $x \vee y=\imposs$ if and only if
$x=y=\imposs$. In relation to the conventional definition, one can
interpret \imposs\ as the closed interval $[0,0]$, \poss\ as the
half-open interval $(0,1]$, and $\vee$ as the addition of
intervals. The definition of a probability measure in this case is
modified to a function $\pmeas : \events \rightarrow \mathscr{L}_{2}$
such that:
\begin{itemize}
\item $\pmeas(\Omega) = \poss$, and
\item for a collection $E_i$ of pairwise disjoint events, $\pmeas(\bigcup_i
E_i) = \bigvee_i \pmeas(E_i)$.
\end{itemize}

@@

\paragraph*{Interval-valued probability measures.} In this case, we
split the $[0,1]$ interval into a collection of overlapping closed
sub-intervals obeying certain restrictions to be described at the end
of the section. First we illustrate the main ideas using a simple
example.

>>>>

We review the conventional presentation of probability spaces and then
discuss several variations that avoid using the real interval $[0,1]$.

%%%
\subsection{Real-Valued Probability Spaces}

@@

\begin{example}[Two-coin probability space]\label{ex1}
  Consider an experiment that tosses two coins. We have four possible
  outcomes that constitute the sample space
  $\Omega = \{ HH, HT, TH, TT \}$. There are 16 total events including
  for example the event $\{ HH, HT \}$ that the first coin lands heads
  up, the event $\{ HT, TH \}$ that the two coins land on opposite
  sides, and the event $\{ HT, TH, TT\}$ that at least one coin lands
  tails up. Here is a possible probability measure for these events:
\[\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) &=& 0 \\
\pmeas(\{ HH \}) &=& 1/3 \\
\pmeas(\{ HT \}) &=& 0 \\
\pmeas(\{ TH \}) &=& 2/3 \\
\pmeas(\{ TT \}) &=& 0 \\
\pmeas(\{  HH, HT \}) &=& 1/3 \\
\pmeas(\{  HH, TH \}) &=& 1 \\
\pmeas(\{  HH , TT \}) &=& 1/3 
\end{array} & \begin{array}{rcl}
\pmeas(\{  HT, TH \}) &=& 2/3 \\
\pmeas(\{  HT , TT \}) &=& 0 \\
\pmeas(\{  TH , TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH \}) &=& 1 \\
\pmeas(\{  HH, HT, TT \}) &=& 1/3 \\
\pmeas(\{  HH, TH, TT \}) &=& 1 \\
\pmeas(\{  HT, TH, TT \}) &=& 2/3 \\
\pmeas(\{  HH, HT, TH, TT \}) &=& 1
\end{array}
  \end{array}\]
  The assignment satisfies the two constraints for probability measures:
  the probability of the entire sample space is 1, and the probability
  of every collection of disjoint events (e.g.,
  $\{ HT \} \cup \{ TH \} = \{ HT, TH \}$) is the sum of the individual
  probabilities. The probability of collections of non-disjoint events
  (e.g., $\{ HT, TH \} \cup \{ TH , TT \} = \{ HT, TH, TT \}$) may add
  to something different than the probabilities of the individual
  events. It is useful to think that this probability measure is
  completely induced by the two coins in question and their
  characteristics in the sense that each pair of coins induces a
  measure, and each measure must correspond to some pair of coins.  The
  measure above is induced by two coins such that the first coin is
  twice as likely to land tails up than heads up and the second coin is
  double-headed.  
  \qed\end{example}

@@

%%%%%
\subsection{Set-valued Probability Measures}

Instead of using every point in the real interval $[0,1]$ we can
partition the interval into disjoint sets and only consider
probability measures up to set membership. The simplest such situation
is to partition the interval $[0,1]$ into $\{0\}$ (which we will call
$\imposs$) and the half-open interval $(0,1]$ (which we will call
$\poss$). The addition that was used to aggregate probabilities is now
abstracted to $\vee$ such that $x \vee y=\imposs$ if and only if
$x=y=\imposs$. We will call the resulting set
$\left\{ \imposs, \poss\right\}$ together with the with associated
operation $\vee$, the set $\mathscr{L}_{2}$. The definition of a
probability measure in this case is modified to a function
$\pmeas : \events \rightarrow \mathscr{L}_{2}$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = \poss$, and
\item for a collection $E_i$ of pairwise disjoint events, $\pmeas(\bigcup_i
E_i) = \bigvee_i \pmeas(E_i)$.
\end{itemize}

@@

%%%%%
\subsection{Interval-valued probability measures}

In this case, we split the $[0,1]$ interval into a collection of
overlapping closed sub-intervals obeying certain restrictions to be
described at the end of the section. First we illustrate the main
ideas using a simple example.
